---
title: "Akshay Jagadeesh, Homework 1"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
---

Load packages you'll be using throughout the HW

```{r}
library(ggplot2)
```

Question 1
-----------------------------------------

### Q1 Part A

```{r 1a}
d0 = c(26, 31, 45)
rs1 = chisq.test(d0, p=c(1, 1, 1), rescale.p=T, simulate.p.value=F)
print(rs1)
```

### Q1 Part B-i

```{r 1b_i}
deaths = c(11, 15, 15, 16, 25, 20)
mean_deaths = mean(deaths)
sd_deaths = sd(deaths)
var_deaths = var(deaths)
print(mean_deaths)
print(sd_deaths)
print(var_deaths)
```

The mean of this sample is `r mean_deaths`, indicating that from 1990 to 1995, `r mean_deaths` juveniles died from maltreatment on average yearly. The standard deviation of these juvenile maltreatment deaths was `r sd_deaths`, and the variance of this sample was `r var_deaths`.

### Q1 Part B-ii

```{r 1b_ii}
se_deaths = sd_deaths / sqrt(length(deaths))
print(se_deaths)
```

The standard error of this sample of juvenile deaths from 1990 to 1995 is `r se_deaths`.

### Q1 Part B-iii

Assuming no trend and  $\bar{x}$ = `r mean_deaths`, we derive the algebraic expression for the estimated number of deaths, *T*, over *k* years by the equation:

T = `r mean_deaths` * k

This is a linear function of $\bar{x}$, which assumes that on average `r mean_deaths` juveniles die of maltreatment each year, so after *k* years, we estimate that the total number of juvenile deaths (T) will equal `r mean_deaths`*k*.

### Q1 Part B-iv

The standard deviation of T will be equal to *k* times the standard deviation of $\bar{x}$.

sd(T) = k * sd(\bar{x}) = k * SE(x) = k * `r se_deaths`

### Q1 Part B-v

```{r 1b_v}
deathDF = data.frame(years=c(1990,1991,1992,1993,1994,1995), deaths)
ggplot(deathDF, aes(x=years, y=deaths)) + 
  geom_point() + geom_smooth(method=lm, colour="red") +
  geom_smooth(method=loess, colour="blue")
```
The red line is the linear regression fit of the data, and the blue line is the Loess curve.

Question 2
-----------------------------------------

### Q2 Part A

```{r 2a}
earlyDeaths = read.csv("earlydeaths.csv")
deathTab <- table(factor(earlyDeaths$cause), factor(earlyDeaths$time))
deathTab <- cbind(deathTab, Total = rowSums(deathTab))
deathTab <- rbind(deathTab, Total = colSums(deathTab))
deathTab
```

### Q2 Part B


### Q2 Part C

Question 9
-----------------------------------------

### Q9 Part A

First, we calculated the mean of the non-normal distribution `D0`. 
  
```{r 9a_D0 mean}
x0 = c(1:3, 5, 7, 9)
p0 = c(.2, .4, .24, .10, .05, .01)
mu0 = sum(x0 * p0) / sum(p0)
print(mu0)
```

The mean of `D0` is 2.66.

**Instructor's Note:** You could also write your narrative text as follows, which will ensure that the number you report is always in sync with the code you wrote. This is a bit more complex, but recommended:

The mean of `D0` is `r mu0`.


### Q9 Part B 

Next, we drew a sample of 1000 observations from the skewed distribution `x0`.
  
```{r 9b_x0_sample}
o1 = sample(x0, 1000, replace=T, prob = p0)
```

Then, we made a histogram of `x0` along with the estimated probability density function.

```{r 9b_histogram, fig.width=7, fig.height=6}
df = data.frame(observations=o1)

ggplot(df, aes(x=observations)) +
  geom_histogram(aes(y=..density..), #..density.. normalizes the y axis to the density (i.e. counts / total)
                 color='black',
                 fill='white',
                 binwidth=1,
                 breaks=c(0.5:9.5)) + # see ?geom_histogram for details
  geom_density(alpha=0.2,
               fill='orange',
               adjust=3) +
  theme_bw()

## If you prefer plotting without ggplot, you can use this code:

# rs1 = hist(o1, prob=T, main='Sampling distribution of x0, 1000 Samples')
# lines(density(o1, adjust=3))
```


Question 10: (example stats writeup)
-----------------------------------------

Here we have some data (`x0`) about the demographics of participants in our experiment. We also have some data (`p0`) about the distribution of races in the United States. 

We want to know whether the distribution of races in our data is consistent with the specified US distribution.
  
```{r 10_create_data}
x0 = c(white=273, black=40, asian=10, other=5)
p0 = c(white=.77, black=.13, asian=.054, other=0.046)

# calculate proportions in sample
x0/sum(x0)
```

#### Conduct a chi-squared test
```{r}
rs = chisq.test(x0, p=p0)
rs

rs$expected
```

Running a $X^2$ goodness-of-fit test, we see that we reject the null hypothesis that the distribution of races in our data is consistent with the US distribution, $X^2$ (`r rs$parameter`, N=`r sum(rs$observed)`) =`r rs$statistic`, p = `r rs$p.value`. Thus, it appears that our data is not consistent with the US distribution. Specifically, we have more white participants than we would have expected (273 (~83%), relative to 253 if 77% of our participants were white), and we have fewer asian and other participants than would have been expected (approximately 10 fewer, off by about 3%). Perhaps we sampled from a college that had a larger proportion of white students relative to the US population, and as a result our sample was biased.
